Probabilistic Automated Language Learning for Configuration Files

### 2016-05-01 OSDI/arch.tex
config: for osdi16

\section{Introduction and System Overview}
\label{sec:Intro}

Configuration errors are one of the most important root-causes of
modern software system failures~\cite{xu15systems,yin11anempirical}.
In practice, misconfiguration problems may result in security
vulnerabilities, application crashes, severe disruptions in software
functionality, and incorrect program executions%
~\cite{xu15systems,zhang14encore,yuan11context}.  Although several
tools have been proposed to automate configuration error diagnosis
after failures occur~\cite{wang04automatic,attariyan10automating,
su07autobash,whitaker04configuration}, these tools rely on a manual approach
  to understand and detect the failure symptoms. The main reasons for this are:
  1) entries in configuration files are untyped assignments, 2) there
  is no explicit structure policy for the entries in configuration
  files, and 3) there are surprisingly few rules specifying the
  entries' constraints.

We propose an approach to the verification of  
configuration files which is based on learning rules about the language 
model for configuration files.

\begin{figure}[t] \centering
\includegraphics[width=0.63\textwidth]{figs/overview}
\caption{{\footnotesize \app's workflow. The green boxes represent configuration files,
  including both correct general configuration files and users' input
  configuration files. The purple boxes are the components within \app.
  The yellow boxes are results generated by \app's components.}}
\label{fig-overview}
\end{figure}

Figure~\ref{fig-overview} describes an overview of our system. We start
with the assumption that we are given a number of correct configuration
files belonging to the same category (for instance, MySQL or
Apache). Such files follow similar patterns, which we exploit
in a learning algorithm to build rules that
describe a language model for the files. Since the
``language'' of configuration types is untyped and unstructured, we
first parse the files and translate them into a more structured,
intermediary representation.
When running type inference on a configuration file, the type of a variable cannot always be fully determined from a single value.
We address this problem by introducing so called {\emph{probabalistic types}}.
Rather than giving a variable a single type, we assign several types with their probability distributions. 
We can then use these more structured files
as a training set to learn the rules. The learning algorithm
is template-based to be easily extensible. We provide an initial set of templates and the
learner learns some concrete instances from the training set. These
rules are used for detecting errors violating the learned constraints
in the files given by the user.

@@

From a practical perspective \app introduces no additional burden 
to the users: they can simply use \app to check for errors in their configuration files. However, they can also easily extend the framework themselves. The system is designed to be highly modular. If there is a class of rules that \app is not currently learning, the user can develop their own templates and learners for that class. The new learner can be added to \app and this way it can check additionally a new set of errors.

Finally, from a systems perspective this is the first approach that {\emph{proactively}} checks 
 the correctness of configuration files. All previous work
~\cite{xu15systems,zhang14encore,yuan11context, wang04automatic,attariyan10automating,
su07autobash,whitaker04configuration} tries to identify the problem after the
failure occured. Our approach isolates potential errors before the system failure occurs, e.g. before the installation. We can also see \app as a tool that can run in conjunction with existing tools. Pre-analyzed configuration files are already free from language-based errors, and this way the workloads of post-failure forensics at the runtime
is significantly reduced, thus making these tools truly practical.

To summarize, this tool paper makes the following contributions:

\begin{enumerate}

  \item We designed and implemented a tool, \app, that can learn a
language model from an example set of correct configuration files, and
we use the model to verify new configuration files.
  \item We use probabilistic types to assign a confidence distribution over a set of types to a value.
  \item In \app we define a interface for describing a verification attribute in a learning context, making it easy to add new rules to the system.

\end{enumerate}

>>>>

\section{\app Overview}

\begin{figure*}[tbp] \centering
\includegraphics[width=0.68\textwidth]{figs/overview}
\caption{{\footnotesize \app's workflow. The green boxes represent configuration files,
  including both correct general configuration files and users' input
  configuration files. The purple boxes are the components within \app.
  The yellow boxes are results generated by \app's components.}}
\label{fig-overview}
\end{figure*}

Figure~\ref{fig-overview} describes an overview of our system. We start
with the assumption that we are given a number of correct configuration
files belonging to the same category, such as MySQL or Apache. 
Such files follow similar patterns, which we exploit
in a learning algorithm to build rules that
describe a language model for the files. Since the
``language'' of configuration types is untyped and unstructured, we
first parse the files and translate them into a more structured,
intermediary representation.
When running type inference on a configuration file, 
the type of a variable cannot always be fully determined from 
a single value.
We address this problem 
by introducing so called {\emph{probabilistic types}}.
Rather than giving a variable a single type, we assign several types with their probability distributions. 
We can then use these more structured files
as a training set to learn the rules. The learning algorithm
is template-based to be easily extensible. We provide an initial set of templates and the
learner learns some concrete instances from the training set. These
rules are used for detecting errors violating the learned constraints
in the files given by the user.

@@

\com{
A rule R is added to the set of all rules,
  if $\exists$ learning file f s.t. R(f) is non vacuously true
That rule R is then removed from the set of all rules,
  if $\exists$ learning file f st R(f) is false.

This is accomplished in two passes.
First we collect all possible rules for every file.
Then we merge the all the rules to create our final set.
This conviently gives rise to an embarresingly parallel situation, which Haskell allows us to easily take advantage of by
  using the parallel mapping library parmap.

\begin{lstlisting}
  potentialRules = parmap findAllRules learningSet.
  finalRules = foldl1 mergeRules potenialRules
\begin{lstlisting}

Each rule is of the Attribute typeclass, which means a rule must support the following operations:
\begin{lstlisting}
class Attribute r where
  learn :: ConfigFile Common -> [r]
  merge :: [r] -> [r] -> [r] 
  check :: [r] -> ConfigFile Common -> Error
\end{lstlisting}

}

### 2016-05-01 OSDI/concl.tex
config: for osdi16

\section{Conclusions}

In this paper, we introduce \app, a highly-modular framework that allows verification of configuration files, even without a language model of the file.
New verification properties require only a small amount of code and are not language specific, all indicating that \app could be widely adopted by system administrators.
Such a verification tool that scales in both performance and expressivity can revolutionize configuration file checking, reducing the cost of system maintaince and failure dramatically.

>>>>

\section{Conclusion}

In this paper, we introduce \app, a highly-modular framework 
that allows verification of configuration files, 
even without a language model of the file.
New verification properties require only a small amount of code 
and are not language specific, 
all indicating that \app could be widely adopted by system administrators.
Such a verification tool that scales in both performance and expressivity 
can revolutionize configuration file checking, 
reducing the cost of system maintenance and failure dramatically.

### 2016-05-01 OSDI/evaluation.tex
config: for osdi16

Since we learn a set of rules on each file in isolation from the other, we have a pleasingly parallel situation.
Haskell allows us to easily take advantage of this by using the parallel mapping library \cite{}, both for translation to the intermediate representation, and for learning the rules on each file.
The merge stage could also easily be parallelized, using a divide and conquer approach, but \app runs fast enough over our learning set (28 files, 961 lines of code) that this has not been necessary.

>>>>

Since we learn a set of rules on each file in isolation from the other, we have a pleasingly parallel situation.
Haskell allows us to easily take advantage of this by using the parallel
mapping library \cite{parallel}, both for translation to the intermediate representation, and for learning the rules on each file.
The merge stage could also easily be parallelized, using a divide and conquer approach, but \app runs fast enough over our learning set (28 files, 961 lines of code) that this has not been necessary.

### 2016-05-01 OSDI/motiv.tex
config: for osdi16

\para{Example~1: Ordering Errors.} When configuring PHP to run with the
Apache HTTP Server, the user writes, among others, the following lines:\\
 \texttt{
 \hspace*{3em}extension = mysql.so\\
 \hspace*{3em}...\\
 \hspace*{3em}extension = recode.so}\\
This file caused the Apache server to fail to start due to a segmentation fault error.
When using PHP in Apache, the extension ``mysql.so'' depends on ``recode.so'' and the relative ordering of two of them is crucial. 
\app would inform the user that ``recode.so'' should appear before ``mysql.so'', and return the error:\\
 \texttt{
ORDERING ERROR: Expected "extension"recode.so"\\
   BEFORE "extension""mysql.so"
  }

\para{Example~2: Entry Missing Errors.} If the user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
done through the following entries in the OpenLDAP configuration file:\\
\texttt{
 \hspace*{3em}include schema/ppolicy.schema\\
 \hspace*{3em}overlay ppolicy\\}
When using the password policy overlay in OpenLDAP, we have to first include the related schema.
Leaving out the ``include'' statement will cause the failure of 
this LDAP. Running \app on such a misconfiguration file would return:\\
\texttt{
MISSING KEYWORD ERROR: Expected "overlay""ppolicy"\\ 
in the same file as: "include""schema/ppolicy.schema"}

\para{Example~3: Type Errors.} If the user tries to install MySQL, she first needs to initiate the path for the log information generated by MySQL. 
A user may put the following code in the MySQL configuration file:\\
\texttt{
 \hspace*{3em}general\_log = /var/log/mysql/mysql.log\\}
However, the entry ``general\_log'' should be an integer, not a string.
 In MySQL, there is another entry named ``general\_log\_file'' which is used to specify the log path.  
After \app analyzes this configuration file, it correctly identifies the error:\\
\texttt{
TYPE ERROR: Expected a Int with P=1.0 for "general\_log[mysqld]"
}

\vspace{-10pt}
\para{Example~4: Value Correlation Errors.} When configuring PHP on MySQL, the user may write the following lines of entries in both the PHP and MySQL configuration files:\\
\texttt{
 \hspace*{3em}mysql's config\\
 \hspace*{3em}max\_connections = 300\\
 \hspace*{3em}...\\
 \hspace*{3em}php's config\\
 \hspace*{3em}mysql.max\_persistent = 400\\}
This could cause MySQL to abort with the error information: ``too many connections''.
In this case, the ``mysql.max\_persistent'' in PHP should be no larger than the ``max\_connections'' in MySQL configuration file.
Another rule we have implemented is learning inequality relations between integers.
Running \app on this combined configuration file would return:
\texttt{
INTEGER RELATION ERROR: \\
Expected "max\_connections">="mysql.max\_persistent"}

>>>>

\para{Example~1: Ordering errors.} 
When configuring PHP to run with the
Apache HTTP Server, the user writes, among others, the following lines:\\
 \texttt{
 \hspace*{3em}extension = mysql.so\\
 \hspace*{3em}...\\
 \hspace*{3em}extension = recode.so}\\
This file caused the Apache server to fail to start due to a segmentation fault error.
When using PHP in Apache, the extension ``mysql.so'' depends on ``recode.so'' and the relative ordering of two of them is crucial. 
\app would inform the user that ``recode.so'' should appear before ``mysql.so'', and return the error:\\
 \texttt{
ORDERING ERROR: Expected "extension"recode.so"\\
   BEFORE "extension""mysql.so"
  }

\para{Example~2: Entry missing errors.} 
If the user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
done through the following entries in the OpenLDAP configuration file:\\
\texttt{
 \hspace*{3em}include schema/ppolicy.schema\\
 \hspace*{3em}overlay ppolicy\\}
When using the password policy overlay in OpenLDAP, we have to first include the related schema.
Leaving out the ``include'' statement will cause the failure of 
this LDAP. Running \app on such a misconfiguration file would return:\\
\texttt{
MISSING KEYWORD ERROR: Expected "overlay""ppolicy"\\ 
in the same file as: "include""schema/ppolicy.schema"}

\para{Example~3: Type errors.} 
If the user tries to install MySQL, she first needs to initiate the path for the log information generated by MySQL. 
A user may put the following code in the MySQL configuration file:\\
\texttt{
 \hspace*{3em}general\_log = /var/log/mysql/mysql.log\\}
However, the entry ``general\_log'' should be an integer, not a string.
 In MySQL, there is another entry named ``general\_log\_file'' which is used to specify the log path.  
After \app analyzes this configuration file, it correctly identifies the error:\\
\texttt{
TYPE ERROR: Expected a Int with P=1.0 for "general\_log[mysqld]"
}

\para{Example~4: Value correlation errors.} 
When configuring PHP on MySQL, the user may write the following lines of entries in both the PHP and MySQL configuration files:\\
\texttt{
 \hspace*{3em}mysql's config\\
 \hspace*{3em}max\_connections = 300\\
 \hspace*{3em}...\\
 \hspace*{3em}php's config\\
 \hspace*{3em}mysql.max\_persistent = 400\\}
This could cause MySQL to abort with the error information: ``too many connections''.
In this case, the ``mysql.max\_persistent'' in PHP should be no larger than the ``max\_connections'' in MySQL configuration file.
Another rule we have implemented is learning inequality relations between integers.
Running \app on this combined configuration file would return:
\texttt{
INTEGER RELATION ERROR: \\
Expected "max\_connections">="mysql.max\_persistent"}
